---
title: "top101 visualization"
author: "MarijnJABoer"
date: "17/01/2018"
output:
  md_document:
    variant: markdown_github
    toc: true
    fig_width: 7
    fig_height: 6
---

```{r Setup, echo=T, message=FALSE, warning=FALSE, results='hide'}
setwd("~/Internships/Internship CNN 2017-2018/FormicID/stat/top101")


library(ggplot2) # for plotting
library(cowplot)
library(jpeg)
library(imager) # for reading jpg files
library(reshape2) # for using melt()
library(magick) # for image conversion because jpg files are read
library(plyr)
library(magrittr)
library(forcats)
library(dplyr)

data_dir <- # Setting the directory that contains all the images
    "~/Internships/Internship CNN 2017-2018/FormicID/data/top97species_Qmed_def_clean"

top101 <-
    read.csv(file.path(data_dir, 'image_urls.csv')) # spreadsheat containing catalognumber, scientific name,
# shot_type and image url
```

# Dataset

```{r small data set inspection}
head(top101)
summary(top101)
str(top101)
species <- data.frame(count(top101$scientific_name))
species[order(species[, 2], decreasing = T), ]
levels(top101$shot_type)[levels(top101$shot_type) == 'd'] <-
    'Dorsal'
levels(top101$shot_type)[levels(top101$shot_type) == 'h'] <- 'Head'
levels(top101$shot_type)[levels(top101$shot_type) == 'p'] <-
    'Profile'

# https://stackoverflow.com/questions/8186436/order-stacked-bar-graph-in-ggplot
top101_transp <-
    read.csv(file.path(data_dir, 'image_urls_transp.csv'), sep = ';')
top101_transp$total <- NULL # remove total
capitalize <- function(x) {
    first <- toupper(substr(x, start = 1, stop = 1))
    rest <- tolower(substr(x, start = 2, stop = nchar(x)))
    paste0(first, rest)
}
levels(top101_transp$scientific_name) <-
    capitalize(levels(top101_transp$scientific_name))
top101_transp[1] <- sapply(top101_transp[1],
                           gsub,
                           pattern = "_",
                           replacement = " ")
top101_transp$scientific_name <-
    reorder(top101_transp$scientific_name, rowSums(top101_transp[2:4])) # reorder based on a non-created total
melted <-
    melt(top101_transp, id = 'scientific_name') # melt based on scientific_name
levels(melted$variable)[levels(melted$variable) == 'dorsal'] <-
    'Dorsal'
levels(melted$variable)[levels(melted$variable) == 'head'] <- 'Head'
levels(melted$variable)[levels(melted$variable) == 'profile'] <-
    'Profile'

shotcount <-
    data.frame(
        Dorsal = sapply(melted[, -1], function(x) {
            sum(melted$value[x == 'Dorsal'])
        }),
        Head = sapply(melted[, -1], function(x) {
            sum(melted$value[x == 'Head'])
        }),
        Profile = sapply(melted[, -1], function(x) {
            sum(melted$value[x == 'Profile'])
        })
    )
shotcount <- melt(shotcount[(1), ])

# Manual fix for the correct numbers 
value = c(3405,3386, 3420) 
variable = c("Dorsal", "Head", "Profile") 
shotcount = data.frame(variable, value) 
```

```{r Data input 2}
data2 <- read.csv("Sup02_top97species_Qmed_def_info.csv", sep = ";")
capitalize <- function(x) {
    first <-
        toupper(substr(x, start = 1, stop = 1)) ## capitalize first letter
    rest <-
        tolower(substr(x, start = 2, stop = nchar(x)))   ## everything else lowercase
    sub("_", " ", paste0(first, rest))
}
levels(data2$scientific_name) <- capitalize(levels(data2$scientific_name))

```


## Shot type plot

```{r Shot type plot, echo = FALSE}
g1 <-
    ggplot(data2, aes(x = shot_type, fill = shot_type)) +
    geom_bar(stat = "count", position = "dodge") +
    coord_cartesian(ylim = c(3300, 3400)) +
    scale_y_continuous(breaks = c(3300, 3350, 3400)) +
    xlab("Shottypes") + ylab("Image count") +
    labs(title = "Shottypes (Clean)") +
    geom_text(
        stat = "count",
        aes(label= ..count..),
        position = position_dodge(width = 0.9),
        vjust = 1.6,
        color = "white"
    ) +
    theme(
        # plot.title = element_text(size = 15),
        # axis.title = element_text(size = 10),
        # legend.title = element_text(size = 10),
        # legend.text = element_text(size = 10),
        legend.position = "none")
g1
```

```{r}
datacaste <- read.csv("Sup02_top97species_Qmed_def_info.csv", sep = ";")
summary(datacaste)
# datacaste <- datacaste[, c(4, 5)]
# freqtable <- table(datacaste$caste)
# castes <-
#     c("Alate queen",
#       "Dealate queen",
#       "Ergatoid male",
#       "Ergatoid queen",
#       "Male",
#       "Worker")
# castes_big <-
#     c("Queens", "Queens", "Males", "Queens", "Males", "Workers")
# caste_df <- as.data.frame.table(freqtable)
# caste_df$Castes <- as.factor(castes_big)
# # levels(caste_df$Var1) <- castes
# caste_df <- melt(caste_df)
# caste_df <- caste_df[order(caste_df$Castes), ]
# caste_df <- caste_df[order(caste_df$value), ]
# caste_df <- caste_df[order(caste_df$Castes), ]
# caste_df$label_ypos <- c(630, NA, NA, 1196, NA, NA)
# sum(caste_df$value)
# head(caste_df)

```

```{r}
castesclean <- read.csv(
    "Sup07_top97species_Qmed_def_clean_castes.csv", sep=";"
)
summary(castesclean)
```

```{r}
# caste_plot <-
#     ggplot(caste_df) +
#     geom_col(aes(
#         x = reorder(Var1, value),
#         y = value,
#         fill = Var1
#     ), position = "dodge") +
#     geom_text(aes(
#         x = Var1,
#         y = value,
#         label = value,
#         vjust = -0.5
#     ),
#     position = position_dodge(width = 0.9)) +
#     # geom_text(aes(x = 0,
#     #               y = 2500,
#     #               label = label_ypos),
#     #           size = 12,
#     #           hjust = -2) +
#     ggtitle("Caste distribution (Unclean)") +
#     xlab("Caste") +
#     ylab("Caste count") +
#     theme(legend.position = "none") +
#     facet_grid(~ Castes, scales = "free_x", space = "free_x")
# caste_plot
# castesclean$caste <- factor(castesclean$caste, levels = c("eM","m", "eQ", "dQ", "aQ", "w"))

datacaste$caste <- ordered(datacaste$caste, levels = c("eM","m", "eQ", "dQ", "aQ", "w"))
levels(datacaste$caste) 
levels(datacaste$caste) <- c("Ergatoid male", "Male", "Ergatoid queen", "Dealate queen", "Alate queen", "Worker")
levels(datacaste$caste_big) <- c("Male", "Queen", "Worker")
# castesclean$caste <- factor(castesclean$caste, levels = c("eM","m", "eQ", "dQ", "aQ", "w"))

caste_plot2 <-
    ggplot(datacaste, aes(x = caste, fill = caste)) +
    geom_bar(stat = "count", position = "dodge") +
    # coord_cartesian(ylim = c(3300, 3400)) +
    # scale_y_continuous(breaks = c(3300, 3350, 3400)) +
    xlab("Caste") + ylab("Image count") +
    labs(title = "Caste distribution") +
    geom_text(
        stat = "count",
        data = filter(
            datacaste,
            caste == "Male" |
                caste == "Dealate queen" |
                caste == "Alate queen" |
                caste == "Ergatoid male" | caste == "Ergatoid queen"
        ),
        aes(x = caste, label = ..count..),
        position = position_dodge(width = 0.9),
        hjust = 0.5,
        vjust = -1.6
    ) +
    geom_text(
        stat = "count",
        data = filter(datacaste, caste == "Worker"),
        aes(x = caste, label = ..count..),
        position = position_dodge(width = 0.9),
        hjust = 0.5,
        vjust = 1.6,
        color = "white"
    ) +
    theme(# plot.title = element_text(size = 15),
        # axis.title = element_text(size = 10),
        # legend.title = element_text(size = 10),
        # legend.text = element_text(size = 10),
        legend.position = "none") +
    facet_grid(~ caste_big, scales = "free_x", space = "free_x")
caste_plot2
```
# Image distribution per species
Here we can see how images are distributed per species. Within species there are 3 different shot types. p = profile, d = dorsal, and h = head view.

```{r Image-species distribution, fig.height = 9, echo = FALSE}
limit_order <- levels(fct_rev(fct_infreq(data2$scientific_name)))
levels(data2$shot_type) <- c("Dorsal", "Head", "Profile")
g2 <-
    ggplot(data2) +
    geom_bar(aes(x = scientific_name, fill = shot_type),
             stat = 'count',
             width = 0.5) +
    labs(title = "Image distribution per species") +
    coord_flip() +
    labs(x = "Species", y = "Image count") +
    ylim(0,250) + 
    geom_text(
        data = subset(data2, shot_type == "Dorsal"),
        stat = "count",
        aes(x = scientific_name, label = ..count..),
        position = position_dodge(width = 0.9),
        hjust = -0.5,
        size=1.5
    ) +
    geom_text(
        data = subset(data2, shot_type == "Head"),
        stat = "count",
        aes(x = scientific_name, label = ..count..),
        position = position_dodge(width = 0.9),
        hjust = -0.5,
        size=1.5
    ) +
        geom_text(
        data = subset(data2, shot_type == "Profile"),
        stat = "count",
        aes(x = scientific_name, label = ..count..),
        position = position_dodge(width = 0.9),
        hjust = -0.5,
        size=1.5
    ) +
    theme(legend.position = "none",
          axis.text = element_text(size = 6.5, face = "italic"))  +
    scale_x_discrete(limits = limit_order) +
    facet_grid(~ shot_type, scales = "free_x", space = "free_x")
g2
# exampledata <- image_read("dataset_example2.png")
```

```{r old code, without caste, but with example images}
# prefinal <- ggdraw() +
#     draw_plot(g2, 0, 0, 1, 1) +
#     draw_image(exampledata, 0.2, -0.22, scale = 0.55) +
#     draw_plot(g1, 0.45, 0.7, 0.45, 0.2) +
#     draw_plot(caste_plot, 0.45, 0.48, 0.5, 0.25) +
#     draw_plot_label(c("A", "B", "C                  Dataset examples", "D"),
#                     c(0, 0.5, 0.25, 0.5),
#                     c(1, 0.8455, 0.5, 0.7),
#                     size = 15)
# title <-
#     ggdraw() + draw_label("Formicidae image dataset (clean)",
#                           size = 18,
#                           fontface = 'bold')
# final <- plot_grid(title,
#                    prefinal,
#                    ncol = 1,
#                    rel_heights = c(0.05, 1))
# # final
```


```{r castes, shottype and species}
final2 <-
    plot_grid(
        g2,
        caste_plot2,
        nrow = 2,
        labels = c("a)", "b)"),
        rel_heights = c(1, 0.4)
    )
final2
# final3 <- ggdraw() +
#     draw_plot(g2, 0, 0, 1, 1) +
#     draw_plot(
#         final2,
#         x = 0.32,
#         y = 0.10,
#         width = 0.585,
#         height = 0.75
#     ) +
#     draw_plot_label("A", 0, 1)
# final3
```


```{r}

save_plot("testspeciescastes2.pdf",
          final2,
          base_width = 8.27,
          base_height = 11.69)
```

```{r for PP, fig.height = 9, echo = FALSE}
g2 <-
    ggplot(melted) +
    geom_bar(aes(
        x = reorder(scientific_name, -value, sum),
        y = value,
        label = value,
        fill = variable
    ),
    stat = 'identity',
    width = 0.5) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.6)) +
    labs(title = "Image-species distribution per shottype") +
    # coord_flip() +
    labs(x = "Species", y = "Image count") +
    theme(legend.position = "top",
          axis.text = element_text(size = 10))  +
    scale_x_discrete(limits = (levels(melted$scientific_name))) +
    scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500, 600)) +
    scale_fill_discrete(name = 'Shottype',
                        labels = c('Dorsal', 'Head', 'Profile'))

final2 <-
    plot_grid(
        g1,
        caste_plot,
        nrow = 1,
        labels = c("B", "C"),
        rel_widths = c(0.6, 1)
    )
final3 <- ggdraw() +
    draw_plot(g2, 0, 0, 1, 1) +
    draw_plot(
        final2,
        x = 0.2,
        y = 0.35,
        width = 0.585,
        height = 0.55
    ) +
    draw_plot_label("A", 0, 1)

final3 <- plot_grid(title,
                    final3,
                    ncol = 1,
                    rel_heights = c(0.05, 1))
final3
```


# File size distribution 
File size is important because it is related to the pixel dimensions of images and therefore it will have an impact on training / validation speed.

```{r File size - reading data, echo = FALSE}
path <-  file.path(
            data_dir,
            "images"
        )
files_info <-
    file.info(list.files(
        path = path,
        pattern = ".jpg",
        full.names = TRUE,
        recursive = TRUE
    ))
head(files_info)
```

## What is the mean size in kilobytes?

```{r File size - convert and mean}
files_info$size2 <-
    files_info$size / 1000 # Convert bytes to kilobytes
size_kb <-
    files_info$size2 # Creating a variable containing only the file size in Kb
# summary(size_kb)
# boxplot(size_kb) 
cat('The mean size is ', mean(size_kb), 'kb') # What is the mean size?
```

## Plotting file size 
File size is viewed in Kb's.

```{r File size - plotting, echo=FALSE}
# Link to a small explanation on plotting histograms and density curves.
# https://www.statmethods.net/graphs/density.html
size_kb_df <- data.frame(size_kb)

g <- ggplot(size_kb_df, aes(size_kb)) +
    geom_histogram(
        binwidth = 1,
        colour = 'black',
        aes(y = ..density..),
        fill = 'deepskyblue3'
    ) +
    stat_function(fun = dnorm,
                  color = 'red',
                  args = list(
                      mean = mean(size_kb_df$size_kb),
                      sd = sd(size_kb_df$size_kb)
                  )) +
    labs(title = "Histogram of file size density") +
    labs(x = "File size (Kb)", y = "Density (%)")  +
    theme(
        # plot.title = element_text(size = 15),
        # plot.subtitle = element_text(size = 10),
        # axis.title = element_text(size = 10),
        # legend.title = element_text(size = 10),
        # legend.text = element_text(size = 10),
        legend.position = 'top') +
    guides(fill = FALSE)
g
```

# Image dimensions
to inspect if all images are the same dimension and if they are RGB or gray-scale.
```{r Image dimensions - Reading data, error = TRUE, echo = FALSE}
# Reading all images paths in to a 'list'
images <-
    list.files(path = data_dir,
               pattern = ".jpg",
               full.names = TRUE,
               recursive = TRUE)
str(images)

x <- image_read(images[1])
str(x)
# Creates an empty list to be filled in the next function
lst2 <- c()

# Returns dimensions, depth, and channels for the images
for (image in images) {
    # This function creates an error, but it works.
    x <- load.image(image)
    x <- dim(x)
    x <- as.numeric(unlist(x))
    lst2 <- c(lst2, x)
}

# Converting to a Dataframe and get rid of images without 3 channels (RGB)
df <-
    data.frame(matrix(
        unlist(lst2),
        nrow = 10118,
        byrow = TRUE,
        ncol = 4
    ))

# Changing columnnames and rownames
colnames(df) <- c('height', 'width', 'depth', 'channel')
rownames(df) <- images

# Removing grayscale images to a different dataframe
df_channel_wrong <- subset(df, channel <= 2)
df_good <- subset(df, channel > 2)
df_good$names <- rownames(df_good)

df_good2 <- as.data.frame(sapply(df_good,gsub,pattern="//fs-smb-019.ad.naturalis.nl/homedir/Marijn.Boer/Internships/Internship CNN 2017-2018/FormicID/data/top97species_Qmed_def_clean/images/dorsal/1-training/amblyopone_australis/", replacement = ""))
df_good2 <- as.data.frame(sapply(df_good2,gsub,pattern=".jpg", replacement = ""))
library(stringr)
df_shot_types <- str_sub(df_good2$names, -1, -1)
df_good2$shot_type <- df_shot_types
df_good_melt <- melt(df_good2[1:6])
summary(df_good_melt)
df_good_melt$height <- as.numeric(as.character(df_good_melt$height))
df_good_melt$width <- as.numeric(as.character(df_good_melt$width))
df_good_melt$shot_type <- as.factor(df_good_melt$shot_type)
df_good_small <- df_good_melt[,c(1,2,6)]
melted_df <- melt(df_good_small)
summary(melted_df)
# keep <- "Width"
# shotypesformean <- filter(df, 
#                 x %in% "Width")
# aggregate(x=shotypesformean$value,
#           by=list(shotypesformean$shottype),
#           FUN=mean)
```

## Plotting the image dimensions

```{r Image dimensions - Plotting, echo=FALSE}
# Boxplot of the hight and width distribution
g <- ggplot(melted_df, aes(x = variable, y = value)) + 
    geom_boxplot(aes(fill = shot_type), alpha = 0.7) +
    scale_y_continuous(name = "Number of Pixels",
                       breaks = seq(0, 500, 25),
                       limits = c(0, 500)) +
    scale_x_discrete(name = "Channel", labels = c('Height', 'Width')) +
    labs(title = "Boxplot of hight and width distribution",
         subtitle = "for FormicID-97") +
    theme(
        plot.title = element_text(size = 15),
        plot.subtitle = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = 'top'
    ) +
    scale_fill_discrete(name = 'Shot type',
                        labels = c('dorsal', 'head', 'profile')) 
g
# histogram of distribution of the width
h <- hist(
    df_good$width,
    breaks = 50,
    col = "red",
    xlab = "Number of pixels",
    main = "Pixel width distribution of the top 101 most imaged species"
)
h
# Plotting a normal distribution over the histogram
xfit <- seq(min(df_good$width), max(df_good$width), length = 40)
yfit <-
    dnorm(xfit,
          mean = mean(df_good$width),
          sd = sd(df_good$width))
yfit <- yfit * diff(h$mids[1:2]) * length(df_good$width)
lines(xfit, yfit, col = "blue", lwd = 2)
```

## Image examples

```{r image examples, echo = FALSE, eval = FALSE}
img_d <- image_read(images[1])
img_h <- image_read(images[2])
img_p <- image_read(images[3])
casent0102125 <- c(img_h, img_d, img_p)
image_info(casent0102125)
a_australis <- image_append(image_scale(casent0102125, "x52"))
image_write(a_australis, path = 'a_australis.png', format = 'png')

top101_nobroken <- top101[-c(387:401,5111:5113),]
unique_species <- top101_nobroken[!duplicated(top101[2:3]),]
unique_species$image_url <- as.character(unique_species$image_url)
str(unique_species)
urllist <- unique_species[,4]

urllist_test <- unique_species[1:25,4]
images_all <- c()
for (row in urllist_test) {
    urllink <- urllist_test
    img <- image_read(urllink)
    images_all <- c(images_all, img)
}
images_all

```
