---
title: "top101 visualization"
author: "MarijnJABoer"
date: "17/01/2018"
output:
  md_document:
    variant: markdown_github
    toc: true
    fig_width: 7
    fig_height: 6
---

```{r Setup, echo=T, message=FALSE, warning=FALSE, results='hide'}
setwd(
    "~/Google Drive/4. Biologie/Studie Biologie/Master Year 2/Internship CNN/8. FormicID/FormicID/stat/top101"
)

library(ggplot2) # for plotting
library(jpeg)
library(imager) # for reading jpg files
library(reshape2) # for using melt()
library(magick) # for image conversion because jpg files are read

top101 <-
    read.csv('top101.csv') # spreadsheat containing catalognumber, scientific name,
# shot_type and image url
```

# Dataset

```{r small data set inspection}
# head(top101)
summary(top101)
str(top101)
top101_melt <- melt(top101)
species <- data.frame(count(top101_melt$scientific_name))
species[order(species[,2],decreasing=T),]
levels(top101$shot_type)[levels(top101$shot_type)=='d'] <- 'dorsal'
levels(top101$shot_type)[levels(top101$shot_type)=='h'] <- 'head'
levels(top101$shot_type)[levels(top101$shot_type)=='p'] <- 'profile'
```

## Shot type plot

```{r Shot type plot, fig.height = 3, echo = FALSE}
g <- ggplot(top101)
g + geom_bar(aes(shot_type), position = 'dodge') +
    coord_cartesian(ylim = c(3350, 3450)) +
    xlab("Shot types") + ylab("Number of images") +
    ggtitle("Shot type distribution")
```

# Image distribution per species
Here we can see how images are distributed per species. Within species there are 3 different shot types. p = profile, d = dorsal, and h = head view.

```{r Image-species distribution, fig.height = 10, echo = FALSE}
g <-
    ggplot(top101) # Ploting image distribution per species per shot type
g + geom_bar(aes(scientific_name, fill = shot_type), width = 0.5) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.6)) +
    labs(title = "Histogram of species distrubition per shottype",
         subtitle = "from FormicID-97") +
    coord_flip() +
    labs(x = "Number of images", y = "Species") +
    theme(
        plot.title = element_text(size = 25),
        plot.subtitle = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.position = 'top'
    )
```

# File size distribution 
File size is important because it is related to the pixel dimensions of images and therefore it will have an impact on training / validation speed.

```{r File size - reading data, echo = FALSE}
data_dir <- # Setting the directory that contains all the images
    "~/Google Drive/4. Biologie/Studie Biologie/Master Year 2/Internship CNN/8. FormicID/FormicID/data/2018-01-17_top101-images"

files_info <- # Reading all the images in the directory
    file.info(list.files(
        path = data_dir,
        pattern = ".jpg",
        full.names = TRUE
    ))
# head(files_info)
```

## What is the mean size in kilobytes?

```{r File size - convert and mean}
files_info$size2 <-
    files_info$size / 1000 # Convert bytes to kilobytes
size_kb <-
    files_info$size2 # Creating a variable containing only the file size in Kb
# summary(size_kb)
# boxplot(size_kb) 
cat('The mean size is ', mean(size_kb), 'kb') # What is the mean size?
```

## Plotting file size 
File size is viewed in Kb's.

```{r File size - plotting, echo=FALSE}
# Link to a small explanation on plotting histograms and density curves.
# https://www.statmethods.net/graphs/density.html
par(mfrow=c(1,2))
# Histogram
h <- hist(
    size_kb,
    breaks = 50,
    col = "red",
    xlab = "File size (kb)",
    main = "File size distribution of the top 97 most imaged species"
)
# Plotting a normal curve on the histogram
xfit <- seq(min(size_kb), max(size_kb), length = 40)
yfit <- dnorm(xfit, mean = mean(size_kb), sd = sd(size_kb))
yfit <- yfit * diff(h$mids[1:2]) * length(size_kb)
lines(xfit, yfit, col = "blue", lwd = 2)

# Kernel density plot
d <- density(files_info$size2)
plot(d, main = "File size distribution of the top 97 most imaged species")
polygon(d, col = "red", border = "blue")
par(mfrow=c(1,1))
```

# Image dimensions
to inspect if all images are the same dimension and if they are RGB or gray-scale.
```{r Image dimensions - Reading data, error=TRUE, echo = FALSE}
# Reading all images paths in to a 'list'
images <-
    list.files(path = data_dir,
               pattern = ".jpg",
               full.names = TRUE)
str(images)

x <- image_read(images[1])
str(x)
# Creates an empty list to be filled in the next function
lst2 <- c()

# Returns dimensions, depth, and channels for the images
for (image in images) {
    # This function creates an error, but it works.
    x <- load.image(image)
    x <- dim(x)
    x <- as.numeric(unlist(x))
    lst2 <- c(lst2, x)
}

# Converting to a Dataframe and get rid of images without 3 channels (RGB)
df <-
    data.frame(matrix(
        unlist(lst2),
        nrow = 10204,
        byrow = TRUE,
        ncol = 4
    ))

# Changing columnnames and rownames
colnames(df) <- c('height', 'width', 'depth', 'channel')
rownames(df) <- images

# Removing grayscale images to a different dataframe
df_channel_wrong <- subset(df, channel <= 2)
df_good <- subset(df, channel > 2)
df_good_melt <- melt(df_good[1:2])
summary(df_good_melt)
```

## Plotting the image dimensions

```{r Image dimensions - Plotting, echo=FALSE}
# Boxplot of the hight and width distribution
g <- ggplot(df_good_melt)
g + geom_boxplot(aes(x = variable, y = value), alpha = 0.7) +
    scale_y_continuous(name = "Number of Pixels",
                       breaks = seq(0, 225, 25),
                       limits = c(0, 225)) +
    scale_x_discrete(name = "Dimensions") +
    ggtitle("Boxplot of hight and width distribution")

# histogram of distribution of the width
h <- hist(
    df_good$width,
    breaks = 50,
    col = "red",
    xlab = "Number of pixels",
    main = "Pixel width distribution of the top 101 most imaged species"
)
# Plotting a normal distribution over the histogram
xfit <- seq(min(df_good$width), max(df_good$width), length = 40)
yfit <-
    dnorm(xfit,
          mean = mean(df_good$width),
          sd = sd(df_good$width))
yfit <- yfit * diff(h$mids[1:2]) * length(df_good$width)
lines(xfit, yfit, col = "blue", lwd = 2)
```


