---
title: "top101 visualization"
author: "MarijnJABoer"
date: "17/01/2018"
output:
  md_document:
    variant: markdown_github
    toc: true
    fig_width: 7
    fig_height: 6
---

```{r Setup, echo=T, message=FALSE, warning=FALSE, results='hide'}
setwd("~/Internships/Internship CNN 2017-2018/FormicID/stat/top101")


library(ggplot2) # for plotting
library(cowplot)
library(jpeg)
library(imager) # for reading jpg files
library(reshape2) # for using melt()
library(magick) # for image conversion because jpg files are read
library(plyr)

data_dir <- # Setting the directory that contains all the images
    "~/Internships/Internship CNN 2017-2018/FormicID/data/top97species_Qmed_def_clean"

top101 <-
    read.csv(file.path(data_dir, 'image_urls.csv')) # spreadsheat containing catalognumber, scientific name,
# shot_type and image url
```

# Dataset

```{r small data set inspection}
head(top101)
summary(top101)
str(top101)
species <- data.frame(count(top101$scientific_name))
species[order(species[, 2], decreasing = T),]
levels(top101$shot_type)[levels(top101$shot_type) == 'd'] <- 'Dorsal'
levels(top101$shot_type)[levels(top101$shot_type) == 'h'] <- 'Head'
levels(top101$shot_type)[levels(top101$shot_type) == 'p'] <- 'Profile'

# https://stackoverflow.com/questions/8186436/order-stacked-bar-graph-in-ggplot
top101_transp <- read.csv(file.path(data_dir, 'image_urls_transp.csv'), sep = ';')
top101_transp$total <- NULL # remove total
capitalize <- function(x) {
    first <- toupper(substr(x, start = 1, stop = 1))
    rest <- tolower(substr(x, start = 2, stop = nchar(x)))
    paste0(first, rest)
}
levels(top101_transp$scientific_name) <-
    capitalize(levels(top101_transp$scientific_name))
top101_transp[1] <- sapply(
        top101_transp[1],
        gsub,
        pattern = "_",
        replacement = " "
    )
top101_transp$scientific_name <-
    reorder(top101_transp$scientific_name, rowSums(top101_transp[2:4])) # reorder based on a non-created total
melted <-
    melt(top101_transp, id = 'scientific_name') # melt based on scientific_name
levels(melted$variable)[levels(melted$variable) == 'dorsal'] <- 'Dorsal'
levels(melted$variable)[levels(melted$variable) == 'head'] <- 'Head'
levels(melted$variable)[levels(melted$variable) == 'profile'] <- 'Profile'

shotcount <- data.frame(Dorsal = sapply(melted[,-1], function(x){sum(melted$value[x == 'Dorsal'])}), 
           Head = sapply(melted[,-1], function(x){sum(melted$value[x == 'Head'])}),
           Profile = sapply(melted[,-1], function(x){sum(melted$value[x == 'Profile'])}))
shotcount <- melt(shotcount[(1),])
```

## Shot type plot

```{r Shot type plot, echo = FALSE}
g1 <- ggplot(shotcount, aes(x=variable, y=value, fill=variable)) + 
    geom_bar(stat="identity", position = "dodge") +
    coord_cartesian(ylim = c(3350, 3450)) +
    xlab("Shot types") + ylab("Number of images") +
    labs(title = "Shottype distribution") +
    geom_text(aes(label = value), position=position_dodge(width=0.9), vjust=1.6, color="white") +
    theme(
        # plot.title = element_text(size = 15),
        # axis.title = element_text(size = 10),
        # legend.title = element_text(size = 10),
        # legend.text = element_text(size = 10),
        legend.position = "none"
    )
    # scale_fill_manual(values=c("#F8766D", "#00C094", "#00B6EB"))
g1
```

# Image distribution per species
Here we can see how images are distributed per species. Within species there are 3 different shot types. p = profile, d = dorsal, and h = head view.

```{r Image-species distribution, fig.height = 9, echo = FALSE}
g2 <-
    ggplot(melted) + # Ploting image distribution per species per shot type 
geom_bar(aes(x = scientific_name, y = value, fill = variable),
             stat = 'identity',
             width = 0.5) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.6)) +
    labs(title = "Species distribition per shot type") +
    coord_flip() +
    labs(x = "Species", y = "Number of images") +
    theme(
        # plot.title = element_text(size = 15),
        # axis.title = element_text(size = 10),
        # legend.title = element_text(size = 10),
        # legend.text = element_text(size = 10),
        legend.position = "top", 
        legend.background = element_rect(linetype="solid", colour= "Black"),
        axis.text=element_text(size=10)
    )  +
    scale_x_discrete(limits = (levels(melted$scientific_name))) + 
    scale_fill_discrete(name = 'Shot type',
                        labels = c('Dorsal', 'Head', 'Profile')) 

exampledata <- image_read("dataset_example.png")

prefinal <- ggdraw() +
    draw_plot(g2, 0, 0, 1, 1) +
    draw_plot(g1, 0.45, 0.1, 0.5, 0.4) +
    draw_plot_label(c("A", "B"), c(0, 0.45), c(1, 0.5), size = 15) +
    draw_image(exampledata, 0.2, 0.2, clip="on", scale = 1)

title <- ggdraw() + draw_label("Distribution of the clean dataset", size = 18, fontface = 'bold')
final <- plot_grid(title,
                   prefinal,
                   ncol = 1,
                   rel_heights = c(0.05, 1))
save_plot("species_shot_dist.pdf",
          final,
          base_width = 8.27,
          base_height = 11.69)
```



# File size distribution 
File size is important because it is related to the pixel dimensions of images and therefore it will have an impact on training / validation speed.

```{r File size - reading data, echo = FALSE}
files_info <- # Reading all the images in the directory
    file.info(list.files(
        path = data_dir,
        pattern = ".jpg",
        full.names = TRUE
    ))
# head(files_info)
```

## What is the mean size in kilobytes?

```{r File size - convert and mean}
files_info$size2 <-
    files_info$size / 1000 # Convert bytes to kilobytes
size_kb <-
    files_info$size2 # Creating a variable containing only the file size in Kb
# summary(size_kb)
# boxplot(size_kb) 
cat('The mean size is ', mean(size_kb), 'kb') # What is the mean size?
```

## Plotting file size 
File size is viewed in Kb's.

```{r File size - plotting, echo=FALSE}
# Link to a small explanation on plotting histograms and density curves.
# https://www.statmethods.net/graphs/density.html
size_kb_df <- data.frame(size_kb)

g <- ggplot(size_kb_df, aes(size_kb))
g + geom_histogram(
    binwidth = 1,
    colour = 'black',
    aes(y = ..density..),
    fill = 'deepskyblue3'
) +
    stat_function(fun = dnorm,
                  color = 'red',
                  args = list(
                      mean = mean(size_kb_df$size_kb),
                      sd = sd(size_kb_df$size_kb)
                  )) +
    labs(title = "Histogram of file size density",
         subtitle = "for FormicID-97") +
    labs(x = "File size (Kb)", y = "Density (%)")  +
    theme(
        plot.title = element_text(size = 15),
        plot.subtitle = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = 'top'
    ) +
    guides(fill = FALSE)
```

# Image dimensions
to inspect if all images are the same dimension and if they are RGB or gray-scale.
```{r Image dimensions - Reading data, error = TRUE, echo = FALSE}
# Reading all images paths in to a 'list'
images <-
    list.files(path = data_dir,
               pattern = ".jpg",
               full.names = TRUE)
str(images)

x <- image_read(images[1])
str(x)
# Creates an empty list to be filled in the next function
lst2 <- c()

# Returns dimensions, depth, and channels for the images
for (image in images) {
    # This function creates an error, but it works.
    x <- load.image(image)
    x <- dim(x)
    x <- as.numeric(unlist(x))
    lst2 <- c(lst2, x)
}

# Converting to a Dataframe and get rid of images without 3 channels (RGB)
df <-
    data.frame(matrix(
        unlist(lst2),
        nrow = 10204,
        byrow = TRUE,
        ncol = 4
    ))

# Changing columnnames and rownames
colnames(df) <- c('height', 'width', 'depth', 'channel')
rownames(df) <- images

# Removing grayscale images to a different dataframe
df_channel_wrong <- subset(df, channel <= 2)
df_good <- subset(df, channel > 2)
df_good$names <- rownames(df_good)

df_good2 <- as.data.frame(sapply(df_good,gsub,pattern="/Users/nijram13/Google Drive/4. Biologie/Studie Biologie/Master Year 2/Internship CNN/8. FormicID/FormicID/data/2018-01-17_top101-images/", replacement = ""))
df_good2 <- as.data.frame(sapply(df_good2,gsub,pattern=".jpg", replacement = ""))
library(stringr)
df_shot_types <- str_sub(df_good2$names, -1, -1)
df_good2$shot_type <- df_shot_types
df_good_melt <- melt(df_good2[1:6])
summary(df_good_melt)
df_good_melt$height <- as.numeric(as.character(df_good_melt$height))
df_good_melt$width <- as.numeric(as.character(df_good_melt$width))
df_good_melt$shot_type <- as.factor(df_good_melt$shot_type)
df_good_small <- df_good_melt[,c(1,2,6)]
melted_df <- melt(df_good_small)

```

## Plotting the image dimensions

```{r Image dimensions - Plotting, echo=FALSE}
# Boxplot of the hight and width distribution
g <- ggplot(melted_df, aes(x = variable, y = value))
g + geom_boxplot(aes(fill = shot_type), alpha = 0.7) +
    scale_y_continuous(name = "Number of Pixels",
                       breaks = seq(0, 225, 25),
                       limits = c(0, 225)) +
    scale_x_discrete(name = "Channel", labels = c('Height', 'Width')) +
    labs(title = "Boxplot of hight and width distribution",
         subtitle = "for FormicID-97") +
    theme(
        plot.title = element_text(size = 15),
        plot.subtitle = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = 'top'
    ) +
    scale_fill_discrete(name = 'Shot type',
                        labels = c('dorsal', 'head', 'profile')) 

# histogram of distribution of the width
h <- hist(
    df_good$width,
    breaks = 50,
    col = "red",
    xlab = "Number of pixels",
    main = "Pixel width distribution of the top 101 most imaged species"
)
# Plotting a normal distribution over the histogram
xfit <- seq(min(df_good$width), max(df_good$width), length = 40)
yfit <-
    dnorm(xfit,
          mean = mean(df_good$width),
          sd = sd(df_good$width))
yfit <- yfit * diff(h$mids[1:2]) * length(df_good$width)
lines(xfit, yfit, col = "blue", lwd = 2)
```

## Image examples

```{r image examples, echo = FALSE, eval = FALSE}
img_d <- image_read(images[1])
img_h <- image_read(images[2])
img_p <- image_read(images[3])
casent0102125 <- c(img_h, img_d, img_p)
image_info(casent0102125)
a_australis <- image_append(image_scale(casent0102125, "x52"))
image_write(a_australis, path = 'a_australis.png', format = 'png')

top101_nobroken <- top101[-c(387:401,5111:5113),]
unique_species <- top101_nobroken[!duplicated(top101[2:3]),]
unique_species$image_url <- as.character(unique_species$image_url)
str(unique_species)
urllist <- unique_species[,4]

urllist_test <- unique_species[1:25,4]
images_all <- c()
for (row in urllist_test) {
    urllink <- urllist_test
    img <- image_read(urllink)
    images_all <- c(images_all, img)
}
images_all

```
