---
title: "Species complexity"
author: "MarijnJABoer"
date: "09/07/2018"
output:
  md_document:
    variant: markdown_github
    toc: true
    fig_width: 15
    fig_height: 6
---

```{r Setup, echo=T, message=FALSE, warning=FALSE, results='hide'}
setwd("~/Internships/Internship CNN 2017-2018/FormicID/stat/hdp")


library(ggplot2) # for plotting
library(cowplot)
library(jpeg)
library(imager) # for reading jpg files
library(reshape2) # for using melt()
library(magick) # for image conversion because jpg files are read
library(plyr)
library(magrittr)
library(PRROC)

```
# Loading data
```{r}
sp_complex <-
    read.csv("speciescomplexexport.csv", sep=";")
sp_complexall <- 
    read.csv("speciescomplexexportall.csv", sep=";")
```

# Function: getting equation

```{r}
lm_eqn = function(df) {
    m = lm(y ~ x, df)
    eq <- substitute(
        italic(y) == a + b %.% italic(x) * ","
        ~~ italic(r) ^ 2 ~ "=" ~ r2,
        list(
            a = format(coef(m)[1], digits = 2),
            b = format(coef(m)[2], digits = 2),
            r2 = format(summary(m)$r.squared, digits = 3)
        )
    )
    as.character(as.expression(eq))
}
```

## Equation
```{r}
eq <- ddply(sp_complex,.(Shottypes),lm_eqn)
eq2 <- ddply(sp_complexall,.(Shottypes),lm_eqn)
```

# F1 Score
F1 Score is needed when you want to seek a balance between Precision and Recall. Right…so what is the difference between F1 Score and Accuracy then? We have previously seen that accuracy can be largely contributed by a large number of True Negatives which in most business circumstances, we do not focus on much whereas False Negative and False Positive usually has business costs (tangible & intangible) thus F1 Score might be a better measure to use if we need to seek a balance between Precision and Recall AND there is an uneven class distribution (large number of Actual Negatives).


# Plotting
## Number of species in complex vs F1-score
```{r}
sp_complex_plot <-
    ggplot(sp_complex, aes(x = x,
                           y = y,
                           col = Shottypes)) +
    geom_point() +
    geom_smooth(method = "lm",
                se = FALSE,
                formula = y ~ x) +
    geom_text(
        data = eq,
        aes(x = 10, y = 0.15, label = V1),
        parse = T,
        inherit.aes = FALSE
    ) +
    facet_grid(~ Shottypes, scales = "free_x", space = "free_x") +
    ggtitle("Species complexity vs F1-score") +
    xlab("Number of species in complex") +
    ylab("F1-score") +
    theme(legend.position = "none")

sp_complex_plot
```

## Support vs F1-Score
```{r}
sp_complexall_plot <-
    ggplot(sp_complexall, aes(x = x,
                           y = y,
                           col = Shottypes)) +
    geom_point() +
    geom_smooth(method = "lm",
                se = FALSE,
                formula = y ~ x) +
    geom_text(
        data = eq2,
        aes(x = 15, y = 0.5, label = V1),
        parse = T,
        inherit.aes = FALSE
    ) +
    xlim(c(0,25)) +
    facet_grid(~ Shottypes, scales = "free_x", space = "free_x") +
    ggtitle("Support vs F1-score") +
    xlab("Support") +
    ylab("F1-score") +
    theme(legend.position = "none")

sp_complexall_plot
```


# PR + ROC curves
I don't think these curves are correct
## Dorsal

```{r}
par(mfrow=c(1,1))
dorsalpr <- 
    read.csv("T97_CaAll_QuM_ShD_AugM_D05_LR0001_E200_I4_def_clean_best_classification_report.csv", sep="," )
plot(dorsalpr$recall, dorsalpr$precision)
# abline(lm(dorsalpr$recall~dorsalpr$precision), col="red") # regression line (y~x) 
lines(lowess(dorsalpr$recall, dorsalpr$precision), col="blue") # lowess line (x,y)


pr <- pr.curve(dorsalpr$recall, dorsalpr$precision)
print(pr)
# compute area under ROC curve
roc <- roc.curve(dorsalpr$recall, dorsalpr$precision)
print( roc )
# compute PR curve and area under curve
pr <- pr.curve(dorsalpr$recall, dorsalpr$precision, curve = TRUE )
# plot curve
plot(pr)
# compute ROC curve and area under curve
roc <- roc.curve(dorsalpr$recall, dorsalpr$precision, curve = TRUE)
# plot curve
plot(roc)
```
## Head
```{r}
headpr <- 
    read.csv("T97_CaAll_QuM_ShH_AugM_D05_LR0001_E200_I4_def_clean_best_classification_report.csv", sep="," )
pr <- pr.curve(headpr$recall, headpr$precision)
print(pr)
roc <- roc.curve(headpr$recall, headpr$precision)
print( roc )
pr <- pr.curve(headpr$recall, headpr$precision, curve = TRUE )
plot(pr)
roc <- roc.curve(headpr$recall, headpr$precision, curve = TRUE)
plot(roc)
```

## Profile
```{r}
profilepr <- 
    read.csv("T97_CaAll_QuM_ShP_AugM_D05_LR0001_E200_I4_def_clean_best_classification_report.csv", sep="," )
pr <- pr.curve(profilepr$recall, profilepr$precision)
print(pr)
roc <- roc.curve(profilepr$recall, profilepr$precision)
print( roc )
pr <- pr.curve(profilepr$recall, profilepr$precision, curve = TRUE )
plot(pr)
roc <- roc.curve(profilepr$recall, profilepr$precision, curve = TRUE)
plot(roc)

```